Using structural and evolutionary information to correct RNA pyrosequencing errors in non-coding RNAs

### 2013-05-24 JCB/conclusion_JCB.tex

\TODOJerome{Etoffer la conclusion}
In this article we presented a new and efficient way of exploring the mutational landscape of an RNA under structural constraints,
and apply our techniques to identify and fix sequencing errors. In addition, we introduced a new scoring scheme to measure the
likelihood of sequencing errors that combines the classical nearest-neighbour energy model parameters \cite{Turner2010} to the recently introduced
isostericity matrices \cite{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring during base pair replacements.
Importantly, our algorithm runs in  $\Theta(n\cdot(|\Omega|+m^2))$ time and $\Theta(n\cdot(|\Omega|+m))$ memory, where $n$ is the length of the RNA,
$m$ the number of mutations and $\Omega$ the size of the multiple sequence alignment. This achievement enables us to envision applications to high-throughput sequencing pipe-lines.

>>>>

\TODOJerome{Etoffer la conclusion}
In this article we presented a new and efficient way of exploring the mutational landscape of an RNA under structural constraints,
and apply our techniques to identify and fix sequencing errors. In addition, we introduced a new scoring scheme to measure the
likelihood of sequencing errors that combines the classical nearest-neighbour energy model parameters \cite{Turner2010} to the recently introduced
isostericity matrices \cite{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring during base pair replacements.
Importantly, our algorithm runs in  $\Theta(n\cdot(|\Omega|+M^2))$ time and $\Theta(n\cdot(|\Omega|+M))$ memory, where $n$ is the length of the RNA,
$M$ the number of mutations and $\Omega$ the size of the multiple sequence alignment. This achievement enables us to envision applications to high-throughput sequencing pipe-lines.

### 2013-05-24 JCB/introduction_JCB.tex

Here, we expand the range of algorithmic techniques previously introduced with \RNAmutants~\cite{Waldispuhl2008}.
Instead of exploring the full conformational landscape and sample mutants, we develop an inside-outside algorithm that enables us
to explore the complete mutational landscape with a \emph{fixed} secondary structure and to calculate exactly mutational probability values. In addition
to a gain into the numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot m^2)$ for the
original version of  \RNAmutants to $\mathcal{O}(n \cdot m^2)$ for \RNApyro, where $n$ is the size of the sequence and $m$ the number of mutations).

@@

We illustrate and benchmark our techniques for point-wise error corrections on the 5S ribosomal RNA. We choose the latter since it has been extensively
used for phylogenetic reconstructions~\cite{Hori1987} and its sequence has been recovered for over 712 species (in the Rfam seed alignment with id
\texttt{RF00001}). Using a leave one out strategy, we perform random distributed mutations on a sequence. While our methodology is restricted to the correction of 
point-wise error in structured regions (i.e. with base pairs), we show that \texttt{RNApyro} can successfully extract a signal that can be used to reconstruct the 
original sequence with an excellent accuracy. This suggests that \RNApyro is a promising algorithm to complement existing tools in the NGS error-correction 
pipeline.

>>>>

Here, we expand the range of algorithmic techniques previously introduced with the \RNAmutants software~\cite{Waldispuhl2008,Waldispuhl2011}.
Instead of exploring the full conformational landscape and sample mutants, we develop an inside-outside algorithm that enables us
to explore the complete mutational landscape with a \emph{fixed} secondary structure and to calculate exactly mutational probability values. In addition
to a gain into the numerical precision, this strategy allows us to drastically reduce the computational complexity ($\mathcal{O}(n^3 \cdot M^2)$ for the
original version of  \RNAmutants to $\mathcal{O}(n \cdot M^2)$ for \RNApyro, where $n$ is the size of the sequence and $M$ the maximal number of mutations).

@@

\TODOTous{Mentionner technologie Illumina 50-75 nts, et modele d'erreur/mutations.}
We illustrate and benchmark our techniques for point-wise error corrections on the 5S ribosomal RNA. We choose the latter since it has been extensively
used for phylogenetic reconstructions~\cite{Hori1987} and its sequence has been recovered for over 712 species (in the Rfam seed alignment with id
\texttt{RF00001}). Using a leave one out strategy, we perform random distributed mutations on a sequence. While our methodology is restricted to the correction of 
point-wise error in structured regions (i.e. with base pairs), we show that \texttt{RNApyro} can successfully extract a signal that can be used to reconstruct the 
original sequence with an excellent accuracy. This suggests that \RNApyro is a promising algorithm to complement existing tools in the NGS error-correction 
pipeline.
